Implications of uniformly distributed, empirically informed priors for phylogeographical model selection: A reply to Hickerson et al.

### 2014-07-29 manuscripts/reply/reply-main.tex
More editing of reply ms.

\change{
Here, we discuss theoretical considerations for using empirically informed
priors for Bayesian model choice.
}
We then evaluate the empirical Bayesian model-averaging approach of
\citet{Hickerson2013} as a potential solution to the biases of \msb revealed by
\citet{Oaks2012}.
In their re-analysis of the dataset of \citet{Oaks2012}, \citet{Hickerson2013}
made an error by mixing different units of time, which makes the results
presented in their response difficult to interpret (see Supporting Information
for details).
We avoid this error, but still find their approach will often spuriously
support clustered divergence models.
Furthermore, the approach provides another means by which the method can
``escape'' models with large parameter space, which manifests in the preference
of models that exclude from consideration the true values of the model's
parameters.
\change{
Our results, show that it is difficult to choose a uniformly distributed prior
on divergence times that is broad enough to confidently contain the true
parameter values while being narrow enough to avoid spurious support for models
with less parameter space.
}
We also explore the predictions of the potential causes of the bias proposed by
\citet{Oaks2012} and \citet{Hickerson2013}.
We show the primary cause of the method's tendency to incorrectly infer shared
divergences is the uniform priors on divergence times leading to small marginal
likelihoods of models with more divergence-time parameters.
We also show that sampling error is not a major contributer to the problem.
Fortunately, \citet{Oaks2014dpp} has introduced a method that uses more
flexible probability distributions to accommodate prior uncertainty in
divergence times without overly inhibiting the marginal likelihoods of models
with more divergence-time parameters, which increases the method's robustness
and power to detect temporal variation in divergences.

@@

\change{
The strong preference for the narrowest prior on divergence times suggests the
approach of \citet{Hickerson2013} is biased toward models with less parameter
space and, as a consequence, will estimate model-averaged posteriors dominated
by models that exclude true values of the model's parameters.
}
We explored this possibility in two ways.
First, we re-analyzed the Philippines dataset using the model-averaging
approach of \citet{Hickerson2013}, but set one of the prior models with a
uniform prior on divergence times that is unrealistically narrow and almost
certainly excludes most, if not all, of the true divergence times of the 22
taxon pairs.
If small likelihoods of large models cause the method to prefer models with
less parameter space, we expect \msb will preferentially sample from this
incorrect model yielding a posterior that is incorrect (i.e., the
model-averaged posterior will be dominated by an incorrect model that excludes
the truth).
Second, we generated simulated datasets for which the divergence times are
drawn from an exponential distribution and applied the approach of
\citet{Hickerson2013} to each of them to see how often the method excludes the
truth.

@@

For our re-analyses of the Philippines dataset we followed the model-averaging
approach of \citet{Hickerson2013}, but with a reduced set of prior models to
avoid their error of mixing units of time (see SI for details).
We used five prior models, all of which had priors on population sizes of
$\meanDescendantTheta{} \sim U(0.0001, 0.1)$ and $\ancestralTheta{} \sim
U(0.0001, 0.05)$.
Following \citet{Hickerson2013}, each of these models had the following
priors on divergence time parameters:
$M_1$, $\divt{} \sim U(0, 0.1)$;
$M_2$, $\divt{} \sim U(0, 1)$;
$M_3$, $\divt{} \sim U(0, 5)$;
$M_4$, $\divt{} \sim U(0, 10)$; and
$M_5$, $\divt{} \sim U(0, 20)$.
We simulated $1\e6$ random samples from each of the models for a total of
$5\e6$ prior samples.
For each model, we retained the 10,000 samples with the smallest Euclidean
distance from the observed summary statistics after standardizing the
statistics using the prior means and standard deviations of the given model.
From the remaining 50,000 samples, we then retained the 10,000 samples with the
smallest Euclidean distance from the observed summary statistics, this time
standardizing the statistics using the prior means and standard deviations
across all five models.
We then repeated this analysis twice, replacing the $M_1$ model with
$M_{1A}$ and $M_{1B}$, which differ only by having priors on divergence
times of $\divt{} \sim U(0, 0.01)$ and $\divt{} \sim U(0, 0.001)$,
respectively.
While we suspect the prior of $\divt{} \sim U(0, 0.1)$ used by
\citet{Hickerson2013} likely excludes the true divergence times of at least
some of the 22 taxa, we are nearly certain that these narrower priors exclude
most, if not all, of the divergence times of the Philippine taxa.

Our results show that the model-averaging approach of \citet{Hickerson2013}
strongly prefers the prior model with the narrowest distribution on divergence
times across all three of our analyses, even when this model is incorrect and
excludes the true divergence times of the Philippine taxa
(Table~\ref{tabModelChoiceEmpirical}).
However, \citet{Hickerson2013} vetted the priors used in their model-averaging
approach via ``graphical checks,'' in which the summary statistics from 1000
random samples of each prior model are plotted along the first two orthogonal
axes of a principle component analysis (see Figure 1 of \citet{Hickerson2013}).
To determine if such prior predictive analyses would indicate the $M_{1A}$ and
$M_{1B}$ models are problematic, we performed these graphical checks on our
prior models.
Unfortunately, these prior predictive checks provide no warning that these
priors are too narrow (Figure~S\ref{figPCA}).
Rather, the graphs suggest these incorrect priors are ``better fit''
(Figure~S\ref{figPCA}A--C) than the valid priors similar to those used by
\citet{Oaks2012} (Figure~S\ref{figPCA}D--F).

@@

For each simulation replicate, we estimated the Bayes factor in favor
of excluding the truth as the ratio of the posterior to prior odds of
excluding the true value of at least one parameter.
Whenever the Bayes factor preferred a model excluding the truth we counted the
number of the 22 true divergence times that were excluded by the preferred
model.
Our results show that the model-averaging approach of \citet{Hickerson2013}
favors a model that excludes the true values of parameters in 97\% of the
replicates (90\% with GLM-regression adjustment), excluding up to 21 of the 22
true divergence times (Figure~\ref{figExclusionSimTau}).
We also used the mode estimate of the preferred model for each replicate to
estimate the number of true parameter values excluded, which produced nearly
identical results.
Importantly, the posterior probability of excluding at least one true parameter
value is very high in nearly all of the replicates
(Figure~\ref{figExclusionSimProb}).
Using a Bayes factor of greater than 10 as a criterion for strong support, 66\%
of the replicates (87\% with GLM-regression adjustment) strongly support the
exclusion of true values (Figure~\ref{figExclusionSimProb}).

@@

\change{
When using uniformly distributed priors, the alternative to priors that reflect
prior uncertainty is to risk excluding the true values one seeks to estimate.
Fortunately, continuous distributions more appropriate as priors for positive
real-valued parameters have been shown to greatly reduce spurious support for
clustered divergence models while allowing prior uncertainty to be accommodated
\citep{Oaks2014dpp}.
}

@@

Our results demonstrate that the approach of \citet{Hickerson2013} consistently
infers highly clustered divergences across all the \divt{max} we simulated
(Figure~\ref{figPower4}A--D \& S\ref{figPower6}A--F).
If we follow \citet{Hickerson2013} and use a Bayes factor of greater than 10 as
the criterion for incorrect inference of a single divergence event, we find
that the approach will often strongly support the extreme case of one
divergence event across all our simulation conditions (Figure~\ref{figPower4}E--H \&
S\ref{figPower6}G--L).
The method also struggles to estimate the variance of divergence times
(\vmratio{}), whether evaluating the unadjusted
(Figure~S\ref{figPowerAccuracy}A--F) or GLM-adjusted
(Figure~S\ref{figPowerAccuracy}G--L) posterior estimates.
Overall, the empirical Bayesian model-averaging approach leads to erroneous
support for highly clustered divergences when populations diverge randomly over
the last $8\globalpopsize$ generations.
To put this on a timescale roughly consistent with a vertebrate mitochondrial
locus (assuming a per-site mutation rate of $1\e{-8}$ per generation) this
translates to 10 million generations.
Assuming a mutation rate consistent with nuclear loci of $1\e{-9}$, this is 100
million generations.

@@

\section{The importance of power analyses to guide applications of \texttt{msBayes}}
\citet{Hickerson2013} presented a power analysis of \msb under a narrow uniform
divergence-time prior of 0--1 coalescent units ago.
They found that under these prior conditions \msb can, assuming a
per-site rate of $1.92\e{-8}$ mutations per generation, detect multiple
divergence events among 18 taxa when the true divergences were random over
hundreds of thousands of generations or more.
\citet{Oaks2012} performed similar power analyses under three uniform
divergence-time priors as narrow as 0--5 coalescent units, and found the method
was able to detect multiple events among 22 taxa when divergences were random
over millions of generations.
It is important that investigators perform power analyses to determine the
method's power for their dataset, and decide if \msb has sufficient temporal
resolution to address their hypotheses; in the case of the Philippines dataset,
it did not.
When doing so, it is important to consider what prior conditions are
relevant to their empirical system.
A divergence-time prior of 0--5 coalescent units is quite narrow considering
that this expresses the prior belief that all 22 taxon pairs diverged within
this window.
Certainly, it is  rare for there to be enough \emph{a priori} information to
be certain that all taxa diverged within the last 4$\globalpopsize$ generations
(i.e., 0--1 coalescent units).
Also, it seems unlikely that when such prior information is available that
being able to detect more than one divergence event in the face of 18 random
divergences over hundreds of thousands of generations will provide much insight
into the evolutionary history of the taxa.
% ; assuming a rate of mutation consistent with nuclear loci ($1\e{-9}$), this
% translates to a temporal resolution of 3 million generations or more.

@@

\section{The causes of support for models of co-divergence}
To determine how best to improve the behavior of \msb, it is important to
determine the mechanism by which broad uniform priors cause incorrect support
for clustered models of divergence.
It is well established that vague priors can be problematic in Bayesian model
selection.
Models that integrate over more parameter space characterized by low
probability of producing the data and relatively high prior density will have
smaller marginal likelihoods \citep{Jeffreys1939,Lindley1957}.
Given the uniformly distributed priors on divergence times employed in \msb,
models with more divergence parameters will be ``averaged'' over much greater
parameter space, all with equal prior weight, and much of it with low
likelihood.
In light of this fundamental statistical issue, it is not surprising that the
method tends to support simple models.

However, \citet{Hickerson2013} conclude that the bias is due to insufficient
sampling.
They argue the widest of the three priors used by \citet{Oaks2012} would
infrequently produce samples with many independent population divergence times
as recent as the estimated gene divergence times presented in \citet{Oaks2012}.
However, this argument assumes the gene divergence times presented in
\citet{Oaks2012} were estimated without error.
These estimates were intended to only provide a rough comparison of the
gene divergence times across the 22 taxa.
These analyses assumed an arbitrary strict per-site mutation rate of $2\e{-8}$
per generation for all taxa, and are, of course, subject to estimation error.
Furthermore, the branch-length units of the gene trees are in millions of
years, whereas the divergence time prior of \msb is in generations, thus
\citet{Hickerson2013} make the implicit assumption that all 22 Philippine taxa
have a generation time of one year.
The argument of \citet{Hickerson2013} that divergence-time estimates of
\citet{Oaks2012} should set an upper bound on the prior for divergence times
seems questionable, especially given our findings presented above regarding the
behavior of \msb when empirically informed uniform priors are employed.

Even if we assume the gene divergence times are estimated without error,
Hickerson et al.'s (\citeyear{Hickerson2013}) argument only applies to one of
the three priors used by \citet{Oaks2012}.
The narrowest prior on divergence times used by \citet{Oaks2012} ($U(0, 5)$)
closely mirrors the range of estimates of gene-divergence times if we assume
the same mutation rate and one generation per year (0--5 million years ago).
Applying Hickerson et al.'s (\citeyear{Hickerson2013}) sampling-probability
argument demonstrates this prior is densely populated with samples with large
numbers of divergence parameters with values younger than the estimated gene
divergence estimates.
Thus, if insufficient prior sampling is to blame for the bias, it should be
much reduced under the narrow prior on \divt{}.
However, the magnitude of the bias is very similar across all three priors
explored by \citet{Oaks2012}.
\citet{Hickerson2013} point out a case where the narrow prior performs
slightly better (panel L of Figures S32, S37, and S38 of \citet{Oaks2012}).
However, it is important to note that these results suffered from a bug
in \msb, and there are many cases after \citet{Oaks2012} corrected the 
bug where the narrow prior performs slightly worse (see panels D--J of
Figures 3 and S12).

To disentangle whether model likelihoods or insufficient prior sampling is to
blame for the method's erroneous support for simple models, we must look at the
different predictions made by these two phenomena.
One example, as discussed by \citet{Oaks2012}, is that insufficient prior
sampling should create large variance among posterior estimates, and thus it
should cause analyses to be highly sensitive to the number of samples drawn
from the prior.
Furthermore, if sampling error is biased toward models with less
parameter space, as suggested by \citet{Hickerson2013}, we expect to see
support for these models decrease as sampling increases.
\citet{Oaks2012} did not see such sensitivity when they compared prior sample
sizes of $2\e6$, $5\e6$, and $10^7$.

@@

As discussed by \citet{Oaks2012}, a straightforward prediction if marginal
likelihoods are causing the preference for simple models is that the bias should
disappear as the model generating the data converges to the prior.
\citet{Oaks2012} tested this prediction by performing 100,000 simulations to
assess the model-choice behavior of \msb when the prior model is correct.
The results confirm the prediction as \msb actually tends to underestimate
the probability of the one-divergence model (see Figure 4 of
\citet{Oaks2012}).
We confirmed this same behavior for the model-averaging approach used by
\citet{Hickerson2013} (see SI text and Figure~S\ref{figValidationMCBehavior}).
These results are not clearly predicted if insufficient sampling was causing
the bias.
Even when the prior is correct, due to the discrete uniform prior on
\numt{} implemented in \msb, models with larger numbers of divergence
events (and thus greater parameter space) will still be less densely
sampled than those with fewer divergence events \citep{Oaks2012}.
Thus, the results of the simulations of \citet{Oaks2012} are more consistent
with the fundamental sensitivity of marginal likelihoods to priors.

This is further demonstrated by the results presented herein that show the
model-averaging approach of \citet{Hickerson2013} prefers models with narrower
\divt{} priors (Table~\ref{tabModelChoiceEmpirical} and
Figs.~\labelcref{figExclusionSimTau,figExclusionSimProb,figPowerExclusion4})
and models with fewer \divt{} parameters (Figure~\ref{figPower4}).
In all of these analyses, each of the prior models have the same number of
samples.
Thus, while sampling error will always be present in any numerical Bayesian
approximation method, insufficient sampling is an unlikely explanation for the
erroneous support for models with less parameter space.
While analyses that sample each model proportional to their parameter space
could be explored, it is clear that the marginal likelihoods under broad
uniform priors on divergence times will be greater for models with fewer
divergence-time dimensions.

% \section{General thoughts on the model of \texttt{msBayes}}
\section{A difficult inference problem}
Our findings are not surprising in light of the difficult inference problem
with which \msb is faced.
% To get a sense of the difficulty of this problem, we tally up all the free
% parameters of the \msb model as applied to the dataset of \citet{Oaks2012}.
% Under the simplest model in \msb (i.e., assuming no migration and no
% intra-locus recombination), the number of parameters for each taxon pair
% include:
% The population sizes of the ancestral and descendant populations
% (\ancestralTheta{}, \descendantTheta{1}{}, \descendantTheta{2}{}),
% the magnitude of population contraction in each of the descendant
% populations (\bottleScalar{1}{} and \bottleScalar{2}{}), the
% timing of these contractions (\bottleTime{}), and the $N-1$ node heights
% (coalescent times) of the gene tree that gave rise to the $N$ gene
% copies sampled from both populations.
% Lastly, there are between one and 22 divergence time parameters \divt{} in
% the vector \divtvector.
When applying the simplest model in \msb (i.e., assuming no migration and no
intra-locus recombination) to the dataset of \citet{Oaks2012} with 22 taxon
pairs, there are 581--602 free parameters that model highly stochastic
coalescent and mutational processes.
Furthermore, under this rich stochastic model, the method is estimating the
probability of 1002 divergence models \citep[i.e., the number of integer
partitions of $Y=22$;][]{Oaks2012}.
Furthermore, all the information in the sequence alignment of each taxon pair
is distilled into four summary statistics:
$\pi$ \citep{Tajima1983}, $\theta_W$ \citep{Watterson1975}, $\pi_{net}$
\citep{Takahata1985}, and $SD(\pi-\theta_W)$ \citep{Tajima1989}.
That gives us a total of 88 summary statistics (four from each of the 22 taxon
pairs), which contain minimal information about many of the $\approx 600$
parameters in the model.
More summary statistics can be used in \msb, but most are highly correlated
with the four default statistics, and thus contribute little additional
information about the parameters from the sequence data.
The large number of parameters and divergence models relative to the amount of
information in the data is undoubtedly a key reason the method lacks robustness
to prior conditions.
% Robustness is an important characteristic of a method to gauge its
% applicability to real-world data, because we know the model and priors will be
% wrong to some degree.

The model-averaging approach of \citet{Hickerson2013} adds an additional
dimension of model choice to sample over eight prior models.
This adds an additional free parameter to the model and, more importantly,
forces the model to sample over 8016 unique models for 22 taxa.
While this approach is trivial to implement, it is a non-trivial extension of
the model.
In theory, the model-averaging approach of \citet{Hickerson2013} is very
appealing.
It leverages a great strength of Bayesian statistical procedures, namely the
ability to obtain marginalized estimates that incorporate uncertainty in
nuisance parameters.
However, when averaging over both narrow-empirical and diffuse uniform priors
for a parameter that is expected to have a very non-uniform likelihood density,
and in the context of a model-choice method that is highly sensitive to priors,
it is not too surprising that the approach struggles.

>>>>

\change{
Here, we discuss theoretical considerations for using empirically informed
priors for Bayesian model choice.
}
We then evaluate the empirical Bayesian model-averaging approach of
\citet{Hickerson2013} as a potential solution to the biases of \msb revealed by
\citet{Oaks2012}.
In their re-analysis of the dataset of \citet{Oaks2012}, \citet{Hickerson2013}
made an error by mixing different units of time, which makes the results
presented in their response difficult to interpret (see Supporting Information
for details).
We avoid this error, but still find their approach will often spuriously
support clustered divergence models.
Furthermore, the approach provides another means by which the method can
``escape'' models with large parameter space, which manifests in the preference
of models that exclude from consideration the true values of the models'
parameters.
\change{
Our results, show that it is difficult to choose a uniformly distributed prior
on divergence times that is broad enough to confidently contain the true
parameter values while being narrow enough to avoid spurious support for models
with less parameter space.
}
We also explore the predictions of the potential causes of the bias proposed by
\citet{Oaks2012} and \citet{Hickerson2013}.
We show the primary cause of the method's tendency to incorrectly infer shared
divergences is the uniform priors on divergence times leading to small marginal
likelihoods of models with more divergence-time parameters.
We also show that sampling error is not a major contributer to the problem.
Fortunately, \citet{Oaks2014dpp} has introduced a method that uses more
flexible probability distributions to accommodate prior uncertainty in
divergence times without overly inhibiting the marginal likelihoods of models
with more divergence-time parameters, which increases the method's robustness
and power to detect temporal variation in divergences.

@@

\change{
The strong preference for the narrowest prior on divergence times suggests the
approach of \citet{Hickerson2013} is biased toward models with less parameter
space and, as a consequence, will estimate model-averaged posteriors dominated
by models that exclude true values of the model's parameters.
}
We explored this possibility in two ways.
First, we re-analyzed the Philippines dataset using the model-averaging
approach of \citet{Hickerson2013}, but set one of the prior models with a
uniform prior on divergence times that is unrealistically narrow and almost
certainly excludes most, if not all, of the true divergence times of the 22
taxon pairs.
If small likelihoods of large models cause the method to prefer models with
less parameter space, we expect \msb will preferentially sample from this
erroneous model yielding a posterior that is incorrect (i.e., the
model-averaged posterior will be dominated by an incorrect model that excludes
the truth).
Second, we generated simulated datasets for which the divergence times are
drawn from an exponential distribution and applied the approach of
\citet{Hickerson2013} to each of them to see how often the method excludes the
truth.

@@

For our re-analyses of the Philippines dataset we followed the model-averaging
approach of \citet{Hickerson2013}, but with a reduced set of prior models to
avoid their error of mixing units of time (see SI for details).
We used five prior models, all of which had priors on population sizes of
$\meanDescendantTheta{} \sim U(0.0001, 0.1)$ and $\ancestralTheta{} \sim
U(0.0001, 0.05)$.
Following \citet{Hickerson2013}, each of these models had the following
priors on divergence-time parameters:
$M_1$, $\divt{} \sim U(0, 0.1)$;
$M_2$, $\divt{} \sim U(0, 1)$;
$M_3$, $\divt{} \sim U(0, 5)$;
$M_4$, $\divt{} \sim U(0, 10)$; and
$M_5$, $\divt{} \sim U(0, 20)$.
We simulated $1\e6$ random samples from each of the models for a total of
$5\e6$ prior samples.
For each model, we retained the 10,000 samples with the smallest Euclidean
distance from the observed summary statistics after standardizing the
statistics using the prior means and standard deviations of the given model.
From the remaining 50,000 samples, we then retained the 10,000 samples with the
smallest Euclidean distance from the observed summary statistics, this time
standardizing the statistics using the prior means and standard deviations
across all five models.
We then repeated this analysis twice, replacing the $M_1$ model with
$M_{1A}$ and $M_{1B}$, which differ only by having priors on divergence
times of $\divt{} \sim U(0, 0.01)$ and $\divt{} \sim U(0, 0.001)$,
respectively.
While we suspect the prior of $\divt{} \sim U(0, 0.1)$ used by
\citet{Hickerson2013} likely excludes the true divergence times of at least
some of the 22 taxa, we are nearly certain that these narrower priors exclude
most, if not all, of the divergence times of the Philippine taxa.

Our results show that the model-averaging approach of \citet{Hickerson2013}
strongly prefers the prior model with the narrowest distribution on divergence
times across all three of our analyses, even when this model is incorrect and
excludes the true divergence times of the Philippine taxa
(Table~\ref{tabModelChoiceEmpirical}).
However, \citet{Hickerson2013} vetted the priors used in their model-averaging
approach via ``graphical checks,'' in which the summary statistics from 1000
random samples of each prior model are plotted along the first two orthogonal
axes of a principle component analysis (see Figure 1 of \citet{Hickerson2013}).
To determine if such prior-predictive analyses would indicate the $M_{1A}$ and
$M_{1B}$ models are problematic, we performed these graphical checks on our
prior models.
Unfortunately, these prior-predictive checks provide no warning that these
priors are too narrow (Figure~S\ref{figPCA}).
Rather, the graphs suggest these incorrect priors are ``better fit''
(Figure~S\ref{figPCA}A--C) than the valid priors similar to those used by
\citet{Oaks2012} (Figure~S\ref{figPCA}D--F).

@@

For each simulation replicate, we estimated the Bayes factor in favor
of excluding the truth as the ratio of the posterior to prior odds of
excluding the true value of at least one parameter.
Whenever the Bayes factor preferred a model excluding the truth we counted the
number of the 22 true divergence times that were excluded by the preferred
model.
Our results show that the model-averaging approach of \citet{Hickerson2013}
favors a model that excludes the true values of parameters in 97\% of the
replicates (90\% with GLM-regression adjustment), excluding up to 21 of the 22
true divergence times (Figure~\ref{figExclusionSimTau}).
We also used the mode estimate of the preferred model for each replicate to
estimate the number of true parameter values excluded, which produced nearly
identical results.
Importantly, the posterior probability of excluding at least one true parameter
value is very high in most replicates
(Figure~\ref{figExclusionSimProb}).
Using a Bayes factor of greater than 10 as a criterion for strong support, 66\%
of the replicates (87\% with GLM-regression adjustment) strongly support the
exclusion of true values (Figure~\ref{figExclusionSimProb}).

@@

\change{
When using uniformly distributed priors, the alternative to capturing prior
uncertainty is to risk excluding the true values one seeks to estimate.
Fortunately, more flexible continuous distributions that are better suited as
priors for the positive real-valued parameters of the \msb model have been
shown to greatly reduce spurious support for clustered divergence models while
allowing prior uncertainty to be accommodated
\citep{Oaks2014dpp}.
}

@@

Our results demonstrate that the approach of \citet{Hickerson2013} consistently
infers highly clustered divergences across all the \divt{max} we simulated
(Figure~\ref{figPower4}A--D \& S\ref{figPower6}A--F).
If we follow \citet{Hickerson2013} and use a Bayes factor of greater than 10 as
the criterion for incorrect inference of a single divergence event, we find
that the approach will often strongly support the extreme case of one
divergence event across all our simulation conditions (Figure~\ref{figPower4}E--H \&
S\ref{figPower6}G--L).
The method also struggles to estimate the variance of divergence times
(\vmratio{}), whether evaluating the unadjusted
(Figure~S\ref{figPowerAccuracy}A--F) or GLM-adjusted
(Figure~S\ref{figPowerAccuracy}G--L) posterior estimates.
Overall, the empirical Bayesian model-averaging approach leads to erroneous
support for highly clustered divergences when populations diverged randomly
over the last $8\globalpopsize$ generations.
For loci with per-site rates of mutation on the order of $1\e{-8}$ and
$1\e{-9}$ per generation, this translates to 10 million and 100 million
generations, respectively.

@@

\section{The importance of power analyses to guide applications of \texttt{msBayes}}
\citet{Hickerson2013} presented a power analysis of \msb under a narrow uniform
divergence-time prior of 0--1 coalescent units ago.
They found that under these prior conditions \msb can, assuming a
per-site rate of $1.92\e{-8}$ mutations per generation, detect multiple
divergence events among 18 taxa when the true divergences were random over
hundreds of thousands of generations or more.
\citet{Oaks2012} performed similar power analyses under three uniform
divergence-time priors as narrow as 0--5 coalescent units, and found the method
was able to detect multiple events among 22 taxa when divergences were random
over millions of generations.
It is important that investigators perform simulations to determine the
method's power for their dataset, and decide if \msb has sufficient temporal
resolution to address their hypotheses; in the case of the Philippines dataset,
it did not.
When doing so, it is important to consider what prior conditions are
relevant to their empirical system.
A divergence-time prior of 0--5 coalescent units is quite narrow considering
that this expresses the prior belief that all 22 taxon pairs diverged within
this window.
Certainly, it is  rare for there to be enough \emph{a priori} information to
be certain that all taxa diverged within the last 4$\globalpopsize$ generations
(i.e., 0--1 coalescent units).
Also, it seems unlikely that when such prior information is available that
being able to detect more than one divergence event in the face of 18 random
divergences over hundreds of thousands of generations will provide much insight
into the evolutionary history of the taxa.
% ; assuming a rate of mutation consistent with nuclear loci ($1\e{-9}$), this
% translates to a temporal resolution of 3 million generations or more.

@@

\section{The causes of support for models of co-divergence}
To determine how best to improve the behavior of \msb, it is important to
determine the mechanism by which broad uniform priors cause incorrect support
for clustered models of divergence.
It is well established that vague priors can be problematic in Bayesian model
selection.
Models that integrate over more parameter space characterized by low
probability of producing the data and relatively high prior density will have
smaller marginal likelihoods \citep{Jeffreys1939,Lindley1957}.
Given the uniformly distributed priors on divergence times employed in \msb,
the likelihood of models with more divergence parameters will be ``averaged''
over much greater parameter space, all with equal prior weight, and much of it
with low likelihood.
In light of this fundamental statistical issue, it is not surprising that the
method tends to support simple models.

However, \citet{Hickerson2013} conclude that the bias is due to insufficient
sampling.
They argue the widest of the three priors used by \citet{Oaks2012} would
infrequently produce samples with many independent population divergence times
as recent as the estimated gene divergence times presented in \citet{Oaks2012}.
However, this argument assumes the gene divergence times presented in
\citet{Oaks2012} were estimated without error.
These estimates were intended to provide only a rough comparison of the
gene divergence times across the 22 taxa.
These analyses assumed an arbitrary strict per-site rate of $2\e{-8}$ mutations
per generation for all taxa, and are, of course, subject to estimation error.
Furthermore, the branch-length units of the gene trees are in millions of
years, whereas the divergence-time prior of \msb is in generations, thus
\citet{Hickerson2013} make the implicit assumption that all 22 Philippine taxa
have a generation time of one year.
The argument of \citet{Hickerson2013} that divergence-time estimates of
\citet{Oaks2012} should set an upper bound on the prior for divergence times
seems questionable, especially given our findings presented above regarding the
behavior of \msb when empirically informed uniform priors are employed.

Even if we assume the arbitrary strict clock is correct, gene divergence times
were estimated without error, and all 22 taxa have one-year generation times,
Hickerson et al.'s
(\citeyear{Hickerson2013}) argument only applies to one of the three priors
used by \citet{Oaks2012}.
Under these assumptions, the narrowest prior on divergence times used by
\citet{Oaks2012} ($U(0, 5)$) closely mirrors the range of estimates of
gene-divergence times (0--5 million years ago).
Applying Hickerson et al.'s (\citeyear{Hickerson2013}) sampling-probability
argument demonstrates this prior is densely populated with samples with large
numbers of divergence parameters with values younger than the estimated gene
divergence estimates.
Thus, if insufficient prior sampling is to blame for the bias, it should be
much reduced under the narrow prior on \divt{}.
However, the magnitude of the bias is very similar across all three priors
explored by \citet{Oaks2012}.
\citet{Hickerson2013} point out a case where the narrow prior performs
slightly better (panel L of Figures S32, S37, and S38 of \citet{Oaks2012}).
However, it is important to note that these results suffered from a bug
in \msb, and there are many cases after \citet{Oaks2012} corrected the 
bug where the narrow prior performs slightly worse (see panels D--J of
Figures 3 and S12).

To disentangle whether model likelihoods or insufficient prior sampling is to
blame for the method's erroneous support for simple models, we must look at the
different predictions made by these two phenomena.
For example, insufficient prior sampling should create large variance among
posterior estimates and cause analyses to be highly sensitive to the number of
samples drawn from the prior.
Furthermore, if sampling error is biased toward models with less
parameter space, as suggested by \citet{Hickerson2013}, we expect to see
support for these models decrease as sampling increases.
\citet{Oaks2012} did not see such sensitivity when they compared prior sample
sizes of $2\e6$, $5\e6$, and $10^7$.

@@

A straightforward prediction if marginal likelihoods are causing the preference
for simple models is that the bias should disappear as the model generating the
data converges to the prior.
\citet{Oaks2012} tested this prediction by performing 100,000 simulations to
assess the model-choice behavior of \msb when the prior model is correct.
The results confirm the prediction as \msb actually tends to underestimate
the probability of the one-divergence model (see Figure 4 of
\citet{Oaks2012}).
We confirmed this same behavior for the model-averaging approach used by
\citet{Hickerson2013} (see SI text and Figure~S\ref{figValidationMCBehavior}).
These results are not clearly predicted if insufficient sampling was causing
the bias.
Even when the prior is correct, due to the discrete uniform prior on the number
of divergence events (\numt{}) implemented in \msb, models with larger numbers
of divergence-time parameters (and thus greater parameter space) will still be
far less densely sampled than those with fewer divergence events
\citep{Oaks2012}.
Thus, the results of the simulations of \citet{Oaks2012} are more consistent
with the fundamental sensitivity of marginal likelihoods to priors.

This is further demonstrated by the results presented herein that show the
model-averaging approach of \citet{Hickerson2013} prefers models with narrower
\divt{} priors (Table~\ref{tabModelChoiceEmpirical} and
Figs.~\labelcref{figExclusionSimTau,figExclusionSimProb,figPowerExclusion4})
and fewer \divt{} parameters (Figure~\ref{figPower4}).
In all of these analyses, each of the prior models have the same number of
samples.
Thus, while sampling error will always be present in any numerical Bayesian
approximation method, insufficient sampling is an unlikely explanation for the
erroneous support for models with less parameter space.
While analyses that sample each model proportional to their parameter space
could be explored, it is clear that the marginal likelihoods under broad
uniform priors on divergence times will be greater for models with fewer
divergence-time dimensions.

% \section{General thoughts on the model of \texttt{msBayes}}
\section{A difficult inference problem}
Our findings are not surprising in light of the difficult inference problem
with which \msb is faced.
% To get a sense of the difficulty of this problem, we tally up all the free
% parameters of the \msb model as applied to the dataset of \citet{Oaks2012}.
% Under the simplest model in \msb (i.e., assuming no migration and no
% intra-locus recombination), the number of parameters for each taxon pair
% include:
% The population sizes of the ancestral and descendant populations
% (\ancestralTheta{}, \descendantTheta{1}{}, \descendantTheta{2}{}),
% the magnitude of population contraction in each of the descendant
% populations (\bottleScalar{1}{} and \bottleScalar{2}{}), the
% timing of these contractions (\bottleTime{}), and the $N-1$ node heights
% (coalescent times) of the gene tree that gave rise to the $N$ gene
% copies sampled from both populations.
% Lastly, there are between one and 22 divergence time parameters \divt{} in
% the vector \divtvector.
When applying \msb to the dataset of \citet{Oaks2012} with 22 taxon
pairs, there are 581--602 free parameters that model highly stochastic
coalescent and mutational processes.
Under this rich stochastic model, the method is estimating the
probability of 1002 divergence models \citep[i.e., the number of integer
partitions of $Y=22$;][]{Oaks2012}.
Furthermore, all the information in the sequence alignment of each taxon pair
is distilled into four summary statistics.
This gives us a total of 88 summary statistics (four from each of the 22 taxon
pairs) that contain minimal information about many of the $\approx 600$
parameters in the model.
More summary statistics can be used in \msb, but most are highly correlated
with the four default statistics, and thus contribute little additional
information about the parameters from the sequence data.
The large number of parameters and divergence models relative to the amount of
information in the data is undoubtedly a key reason the method lacks robustness
to prior conditions.
% Robustness is an important characteristic of a method to gauge its
% applicability to real-world data, because we know the model and priors will be
% wrong to some degree.

The model-averaging approach of \citet{Hickerson2013} adds an additional
dimension of model choice to sample over eight prior models.
This adds an additional free parameter to the model and, more importantly,
forces the model to sample over 8016 unique models for 22 taxa.
While this approach is trivial to implement, it is a non-trivial extension of
the model.
In theory, the model-averaging approach of \citet{Hickerson2013} is very
appealing.
It leverages a great strength of Bayesian statistical procedures, namely the
ability to obtain marginalized estimates that incorporate uncertainty in
nuisance parameters.
However, when averaging over both narrow-empirical and diffuse uniform priors
for a parameter that is expected to have a very non-uniform likelihood density,
and in the context of a model-choice method that is highly sensitive to priors,
it is not surprising that the approach struggles.

The recommendations of \citet{Oaks2012} for mitigating the lack of robustness
of \msb are similar to those of \citet{Hickerson2013}, but avoid the
need for imposing an additional dimension of model choice.
\citet{Oaks2012} suggest that uniform priors may not be ideal for many
parameters of the \msb model, and recommend the use of probability
distributions from the exponential family.
If we look at the prior distribution on divergence-time parameters imposed by
the model-averaging approach of \citet{Hickerson2013} we see it is a mixture of
overlapping uniforms with lower limits of zero (Figure~S\ref{figMCTauPrior}).
This looks very much like an exponential distribution, except that in any state
of the model, all the divergence-time parameters are restricted to the hard
bounds of one of the uniform distributions.
Thus, it seems more appropriate to simply place a gamma prior (the exponential
being a special case) on divergence times.
This would capture the prior uncertainty that \citet{Hickerson2013} are
suggesting for divergence times (Figure~S\ref{figMCTauPrior}) while avoiding
costly model-averaging and the constraint that all divergence times must fall
within the hard bounds of the current model state.
It also would allow an investigator to place the majority of the prior density
in regions of parameter space they believe, \emph{a priori}, are most
plausible, but still capture uncertainty in the tails of distributions with low
density.
Indeed, \citet{Oaks2014dpp} has shown that the use of more flexible
distributions in place of uniform priors improves the power of the method to
detect temporal variation in divergences and reduces incorrect support for
clustered divergences.
